---
title: "Random Intercept models"
author: "Chris Mainey"
date: "19/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a surface-level overview of fitting random-intercept models in R and Python and trying to use them for the prediction task described in our chat on 19/01/2022.  Hopefully it's a useful starter!

For prediction, I will demonstrate two methods for a random-intercept:

* __"Conditional"__ - 'conditioned' on the random-effect i.e. using the random effect.  This gives 'cluster-specific' predictions and in SHMI this would be trust-specific predictions.  You can't use these for a funnel plot, as there is no residual variation and all points line up at 1 on the y-axis (you are summing residual variance in the same clusters you are calculating it at).

* __"Marginal"__ - using the global average prediction i.e. without the random effect.  This gives a global prediction and in SHMI this would be prediction at the national average risk for a patient with a set of predictors (not trust-specific).  You can use these for a funnel plot, you've just got a better case mix model.

Although I advocate the marginal prediction, another approach entirely would be to estimate the random-intercept (how much the trust differs from national average), bootstrap a confidence interval and present as a caterpillar plot.  That's another argument though


## Data

I'm Hilbe's `COUNT` package and the `medpar` dataset which is a cut from 1991 Medicare files for the state of Arizona.

```{r}
library(COUNT)
library(lme4)
library(ModelMetrics)
library(ggplot2)
library(FunnelPlotR)

data("medpar")
```

## In R

This is using the lme4 library which is a frequentest take on multi-level modelling, but it can generally be interpreted in a Bayesian fashion as well, and many mixed-effects model packages are explicitly Bayesian.

### Single-level glm:
```{r rglm}
mod1 <- glm(died ~ age80 + los + factor(type), data=medpar, family="binomial")

summary(mod1)

auc(mod1)
```

### Random-intercept:
```{r rglmer}
mod2 <- glmer(died ~ (1|provnum) + age80 + los + factor(type), data=medpar, family="binomial")

summary(mod2)

auc(mod2)
```

Slightly lower AIC (and asymptotically a reduction of >=4 is a 95% significance, so it's slightly better bad model, slightly improved C-statistic (but it's on the training set).  C-statistics is still rubbish here though.

### Prediction
Remember we are predicting back on to the training set here, so it's better to describe them as 'fitted' I suppose, but it's still the predict function.

```{r rpreds}
# Conditional (cluster-specific)  -  the default
medpar$cond_preds <- predict(mod2, newdata= medpar, type="response")

# Marginal 
medpar$marg_preds <- predict(mod2, newdata= medpar, type="response", re.form = ~0)
```

```{r rpredsout}
head(medpar[c("cond_preds", "marg_preds")])
```
```{r}
ggplot(medpar, aes(y=cond_preds, x=marg_preds, col = as.factor(died)))+
  geom_point()+
  geom_abline(intercept=0, slope=1, col="blue")+
  scale_color_brewer("Died",palette = "Set2")+
  labs(title = "Conditional vs Marginal predictions example in R",
       subtitle = "Blue line: x=y",
       x = "Marginal Prediction (no random-intercept)",
       y = "Conditional Prediciton (with random-intercept)")+
  theme_minimal()+
  theme(legend.position = "bottom")
```

### SHMI equivalent

The equivalent conditional and marginal SMRs would be:

```{r rsmr}
library(dplyr)

SMR_set <-
  medpar %>% 
  group_by(provnum) %>% 
  summarise(num = sum(died),
            conditional = sum(cond_preds),
            marginal = sum(marg_preds),
            SMR_conditional = sum(died) / sum(cond_preds),
            SMR_marginal = sum(died) / sum(marg_preds)
          )

```

### Now Funnel Plot output

#### Conditional
```{r rfunnelcond}
fp1 <- funnel_plot(SMR_set$num, SMR_set$conditional, SMR_set$provnum,limit = 95, draw_adjusted = TRUE)

plot(fp1)
phi(fp1)
```

#### Marginal
````{r funnelplotmarg}
fp2 <- funnel_plot(SMR_set$num, SMR_set$marginal, SMR_set$provnum,limit = 95, draw_adjusted = TRUE)

plot(fp2)
phi(fp2)
```